\propertitle

### 2019-11-18 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-18T20:56:20+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The conclusions regarding \enquote{cooperativity} or \enquote{interaction}
that we reach by applying the maximum-entropy method to the full network,
as proposed here, are not only different from the application at the sample
level, but also visually clearer. The plot of \fig~\ref{fig:plots_2_4mom},
showing the two clearly different distributions constrained by two and four
moments, should be compared with the plot for the distributions obtained at
the sample level under the same constraints, shown in
\fig~\ref{fig:comparisons_marginals} as dashed lines with empty red
circles (\textcolor{myred}{\LARGE$\bm{\mathord{\circ}}$}) and empty blue
squares (\textcolor{mypurpleblue}{$\bm{\square}$}). The measured
frequencies are shown as filled black triangles
(\textcolor{black}{$\bm{\triangledown}$}). It is necessary to use a
logarithmic scale to see the differences, which appear mainly in the tail.
\begin{figure}[!b]
\centering
% \includegraphics[width=\linewidth]{_scripts/comparison_full.pdf}%
% \\[3em]%
\includegraphics[width=\linewidth]{_scripts/trunct_comparison_marginals.pdf}%
\caption{***}
\label{fig:comparisons_marginals}
\end{figure}
% comparison_ME_sample_full_levels.nb

@@

The crucial point is formula~\eqref{eq:equal_m_factorialmoments}, on which
our method is based, and which asserts the equality of the factorial
moments of the full network and those of the sample. This formula is only
valid if our beliefs about the activity of the sample $\pxx(\ya)$ and that
of the full network $\pxx(\yA)$ are related by a hypergeometric
distribution:
\begin{equation}
  \label{eq:hypergeometric_relation}
    \pxx(\ya)=
  \binom{n}{\ya}\;\binom{N-n}{\yA-\ya}\;\binom{N}{\yA}^{-1}
\;\pxx(\yA),
\end{equation}
characteristic of \enquote{drawing without replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
In other words, $N$ must be such that we equally believe other samples of
size $n$ could have been recorded by the electrodes or the probe used. This
is indeed what we stated in \sect~\ref{sec:model}.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-18T22:24:37+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The conclusions about \enquote{cooperativity} or \enquote{interaction}
that we reach by applying the maximum-entropy method to the full network,
as proposed here, are not only different from the application at the sample
level, but also visually clearer. The plot of \fig~\ref{fig:plots_2_4mom},
showing the two clearly different distributions constrained by two and four
moments, should be compared with the plot for the distributions obtained at
the sample level under the same constraints, shown in
\fig~\ref{fig:comparisons_marginals} as dashed lines with empty red
circles (\textcolor{myred}{\LARGE$\bm{\mathord{\circ}}$}) and empty blue
squares (\textcolor{mypurpleblue}{$\bm{\square}$}). The measured
frequencies are shown as filled black triangles
(\textcolor{black}{$\bm{\triangledown}$}). It is necessary to use a
logarithmic scale to see the differences, which appear mainly in the tail.
\begin{figure}[!b]
\centering
% \includegraphics[width=\linewidth]{_scripts/comparison_full.pdf}%
% \\[3em]%
\includegraphics[width=\linewidth]{_scripts/trunct_comparison_marginals.pdf}%
\caption{***}
\label{fig:comparisons_marginals}
\end{figure}
% comparison_ME_sample_full_levels.nb

@@

The crucial point is formula~\eqref{eq:equal_m_factorialmoments}, on which
our method is based, and which asserts the equality of the factorial
moments of the full network and those of the sample. This formula is only
valid if our beliefs about the activity of the sample $\pxx(\ya)$ and that
of the full network $\pxx(\yA)$ are related by a hypergeometric
distribution:
\begin{equation}
  \label{eq:hypergeometric_relation}
    \pxx(\ya)=
  \binom{n}{\ya}\;\binom{N-n}{\yA-\ya}\;\binom{N}{\yA}^{-1}
\;\pxxx(\yA),
\end{equation}
characteristic of \enquote{drawing without replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
In other words, $N$ is the size of the pool of all neurons that could just
as well have been recorded by our measuring procedure. \enquote{Just as
  well} indicates equal degrees of belief. Whether the equality of such
beliefs is justified, or for which sampling procedures it could be
justified, is the fundamental, deep question of sampling theory, which we
cannot discuss here of course \citep[see \eg\ the discussions and
references in][]{ericson1969,smith1976}.

must be such that we equally believe other samples of
size $n$ could have been recorded by the electrodes or the probe used. This
is indeed what we stated in \sect~\ref{sec:model}.
